### 2014-11-02 thesis/litreview/litreview.tex
Fix broken table and figure labels in latex

\chapter{Predicting the LOS: A Review} \label{chap:litreview}

@@

\subsection{Artificial Neural Networks}
Artificial neural networks have been widely used in clinical prediction
following their use in diagnosing acute myocardial infarction and various
applications in radiology \citep{Baxt1995}.
One of the earliest applications of neural networks to
length of stay prediction was by Tu \& Guerriere, who used a neural network
with the following structure \citep{Tu1993} to predict length of stay in
intensive care following cardiac surgery:
\begin{figure}[h]
\caption{}
\centering
\includegraphics[scale=0.95]{images/a2-nn}
\end{figure}

@@

Yang et al. investigated the performance of linear regression, M5 (another
tree-based classification method), and support vector machines in predicting
length of stay for burn patients. The difference between their tree method,
M5, with that of Harper (CART), is that the nodes contain linear regression
functions instead of the usual discrete (two-class) outputs that are used to
successively split the data set. Additionally, they also use support vector
machines (SVMs), perhaps the most powerful classification algorithm in terms of
predictive accuracy \citep{Bellazzi2008}. SVMs work on the principle that when
the data is represented in a high enough dimensional space, a linear hyperplane
can be found where the margin between two different records is as large as
possible. In figure \ref{svm} from Yoo et al. \citep{Yoo2012}, hyperplane 2
is the better plane because its margin is as large as possible. 
\begin{figure}[h]
\label{svm}
\caption{}
\centering
\includegraphics[scale=1.05]{images/a2-svm}
\end{figure}

>>>>

\chapter{Predicting the Length of Stay: A Review} \label{chap:litreview}

@@

\subsection{Artificial Neural Networks}
Artificial neural networks have been widely used in clinical prediction
following their use in diagnosing acute myocardial infarction and various
applications in radiology \citep{Baxt1995}.
One of the earliest applications of neural networks to
length of stay prediction was by Tu \& Guerriere, who used a neural network
with the following structure \citep{Tu1993} to predict length of stay in
intensive care following cardiac surgery:
\begin{figure}[h]
\caption{A graphical representation of a neural network with one hidden layer,
used by Tu \& Guerriere in predicting LOS.}
\centering
\includegraphics[scale=0.95]{images/a2-nn}
\end{figure}

@@

Yang et al. investigated the performance of linear regression, M5 (another
tree-based classification method), and support vector machines in predicting
length of stay for burn patients. The difference between their tree method,
M5, with that of Harper (CART), is that the nodes contain linear regression
functions instead of the usual discrete (two-class) outputs that are used to
successively split the data set. Additionally, they also use support vector
machines (SVMs), perhaps the most powerful classification algorithm in terms of
predictive accuracy \citep{Bellazzi2008}. SVMs work on the principle that when
the data is represented in a high enough dimensional space, a linear hyperplane
can be found where the margin between two different records is as large as
possible. In Figure \ref{fig:svm-example} from Yoo et al. \citep{Yoo2012},
hyperplane 2 is the better plane because its margin is as large as possible. 
\begin{figure}[h]
\caption{An example of finding the largest margin hyperplane for a two-class
support vector machine classifier.}
\label{fig:svm-example}
\centering
\includegraphics[scale=1.05]{images/a2-svm}
\end{figure}

### 2014-11-02 thesis/method/classification.tex
Fix broken table and figure labels in latex

\chapter{LOS Classification Using Supervised Learning}
  \label{chap:classification}

@@

\section{0R and 1R}
\label{sec:zeror}
Short for ``0-rule'' and ``1-rule'' respectively, 0R and 1R are
very simple classifiers. 0R uses no features to predict the class, and
1R only uses one.

Given a data set, the 0R classification algorithm will find the class that
occurs the most often, and predict this class for all new examples. For LOS
classification, if one class in a data set was ``LOS $\leq$ 2 days'' and
the other was
``LOS $>$ 2 days'', and feature vectors (which represent patients in this
scenario) were more commonly labelled ``LOS $\leq$ 2 days'', then 0R will
predict all unseen patients with ``LOS $\leq$ 2 days''.
The prediction decision of 0R is very easy to understand,
and its results also provide an apt baseline for evaluating the more
sophisticated classifiers used in this work.

1R works by considering each feature individually.
For each of the values of the
feature, 1R finds the class with the highest frequency and creates a rule
from this observation. If the feature is numeric (that is, its values are
continuous rather than discrete), then 1R will partition the spectrum of
values into discrete buckets. This is known as \textit{discretisation}.
The error rate, expressed as the ratio of correct predictions to the total
number of examples in the data set,
is calculated for each rule that is generated
in this way. The rule with the lowest error rate is then chosen as the 1-rule
for this classifier,
and if there is a tie, it is broken randomly. As with 0R, the prediction
decision of 1R is transparent and it provides another baseline upon which
more sophisticated classifiers can be evaluated. Studies have also found that
1R performs quite well in terms of accuracy relative to other classifiers
\cite{Holte1993}.

Concretely, given a feature
vector of the form \textit{(age, gender, has\_cancer)},
1R will split the ages of
all patients into discrete categories
(such as \textit{0-20, 21-40, 41-65, 66+}). It
will then check the class labels for all training examples in the data set
that have \textit{age} in the 0-20 category
(as that is the first value for the discretised \textit{age} feature) and
note the most common class, proceeding similarly for the rest of the
categories of \textit{age}. Then it will consider \textit{gender}, which has
two possible values $\{$\textit{male, female}$\}$, and finally \textit{has\_cancer}
with two values $\{$\textit{true, false}$\}$. This results in 8 rules,
corresponding to three features.
An example of a rule is that \textit{age 0-20} most commonly
has ``LOS $\leq$ 2 days''. For each of these rules, 1R will attempt to
use only that rule to predict the LOS class for each example in the data set,
noting down the error rate. 1R then chooses the feature that produces rules
with the lowest total error rate.

\section{Na\"{i}ve Bayes (NB)}
Unlike 1R, which considers only one feature in the feature vector, Na\"{i}ve
Bayes (NB) classifiers allows all features to contribute equally to the
classification decision. Such classifiers apply Bayes' theorem with the
(unrealistic) assumption of independence between features, hence the
description \textit{na\"{i}ve}. This means that we
calculate the probability that a particular feature value belongs in a class
without considering the effect of other features on this probability. This
probability can be computed by
counting the frequency with which the value is associated with a class in the
training data. After these probabilities are computed, classifying a feature
vector simply involves multiplying the probabilities of each feature belonging
to each possible class. The predicted class for that feature vector is then the
class with the highest probability.

@@

Despite the na\"{i}ve independence assumption of the NB classifier, it often
works surprisingly well in practice and is relatively fast to train. We use it
in our work both as a simple classifier in its own right, and also as a more
sophisticated baseline to 0R and 1R.

\section{C4.5 Decision Tree (DT)}
\label{sec:c45}
Decision trees (DT) can be specified recursively and are an example of a
``divide-and-conquer'' approach to learning. First, we choose a feature to
place at the root node and then create one branch for each possible value.
This partitions the data set into subsets, one for every value that the
feature takes. We repeat this for each branch, choosing different features
each time and continuing to divide the data set into progressively smaller
subsets, growing the tree.
If all examples in a subset have the same class at any point during
this process, we stop splitting that subset.

@@

\section{Logistic Regression (LR)}
Logistic regression is a regression technique used for when the dependent
variable is dichotomous or two-valued, and thus it is a suitable candidate to
use for our LOS classification problem.
Previous work on trauma LOS prediction by
Dinh et. al. \cite{Dinh2013a}, as well as many attempts at modelling the LOS
in various medical domains, have all used logistic regression, often as a
starting point.

@@

Training the logistic regression classifier, then, is a matter of finding the
$\beta_i$s that will maximise the likelihood (conditional probability of the
data given the values for $\beta_i$) of the given data set. This is done using
numerical methods, where estimates of the $\beta_i$s are chosen, the likelihood
computed, and the $\beta_i$ are updated, and so on: this process is repeated
until parameter estimates converge or do not change more than a given threshold
value.

To classify a feature vector $x = (x_1,x_2,\ldots,x_n)$, we substitute the
$x_i$'s into the above equation in place of the $F_i$'s, which gives us a
probability. If this probability is above a particular threshold, then we
classify the example as belonging to the class ``LOS $\leq$ 2 days'', otherwise
it is classified as ``LOS $>$ 2 days''. Usually this threshold is 0.5.

\section{Support Vector Machine (SVM)}
Support vector machines (SVMs) are an example of a supervised learning algorithm
capable of classifying training examples into two classes. They are one of the
best algorithms in terms of predictive accuracy \cite{Bellazzi2008}. Despite
this, they have not been widely applied to predicting the LOS, which gives us
an opportunity to evaluate their performance in our LOS classification problem.

@@

\begin{figure}[h]
\label{fig:svm}
\caption{}
\centering
%\includegraphics[scale=0.95]{images/a2-nn}
\end{figure}

@@

\section{$k$-Nearest Neighbour ($k$-NN)}
In this section we will describe the $k$-nearest neighbour ($k$-NN) algorithm
and discuss some extensions which we use in our work. Additionally, we will
describe our contribution, which is an extension of the distance metric for
finding the nearest neighbours.

@@

\subsection{Extensions}
There have been many extensions proposed to the $k$-NN algorithm described
above. These usually seek to find a more suitable distance metric for the
problem domain at hand, or to reduce the storage requirements of the original
$k$-NN algorithm. We will focus on two extensions to the Euclidean distance:
K* and our contribution, Ranked Distance.

@@

\subsubsection{Ranked Distance}
\label{sec:rankeddist}
\paragraph{Intuition}
As we mentioned earlier, all features are considered equally in a $k$-NN
classifier, which implies that all features are equally important in predicting
the class. In practice, however, this is not always the case: to predict the
LOS of a trauma patient, the severity of their injury should affect the final
LOS more than whether or not they can speak English (which could be just two
of many features that are recorded about a patient). Our contribution is thus
a modified distance function to take the relative importance of attributes into
account.

\paragraph{Mathematical formulation}
Given two instances $x = (x_1,x_2,\ldots,x_n)$ and $y = (y_1,y_2,\ldots,y_n)$,
the distance between them is given by:
\begin{equation*}
\mathrm{Distance} = \sum_{i=1}^n w_i |x_i-y_i|
\end{equation*}
where the $w_i$s weight the contribution of the $i$-th feature to the overall
distance that is computed by the $k$-NN algorithm. We assume that all feature
values have been normalised, as described above.

@@

\noindent We will use $f(\mathrm{Rank}) = \frac{1}{\mathrm{Rank}}$ and
$f(\mathrm{Rank}) = n-\mathrm{Rank}$.
\end{enumerate}

\section{Multi-layer Perceptron (MLP)}
Multi-layer perceptrons (MLPs) are a generalised form of the
\textit{perceptron}, first
studied extensively by Rosenblatt \cite{Rosenblatt1962}. The original
motivation of the perceptron 
was to develop a very simple artificial neural network to model the
behaviour of neurons in the brain. The perceptron model for a two-class
problem is a linear combination of the features: that is,
$h(\mathbf{x}) = \sum w_i x_i$ where the $w_i$ are the weights of the model.
There is an input node, called a \textit{neuron}, for each feature, and the
weights are applied to the connections between these input neurons to the
output.
Usually, a constant term of 1 called the \textit{bias} is also added to the
model for an additional trainable input. To classify an item, we substitute it
into $h(\mathbf{x})$; if the value is above a particular threshold, we label it
as class 1, else it belongs in class 2. For convenience, we can let this
threshold be 0. Training the perceptron involves the following steps:
\begin{enumerate}
\item Start with random numbers for the weights $w_i$.
\item For each training example $\mathbf{x_i}$: calculate $h(\mathbf{x_i})$.
If it is correctly classified, do not update the weights.
If it is incorrectly classified, add a multiple of the misclassified
instance to the current weights using
$\mathbf{w} = \mathbf{w} + \lambda \mathbf{x_i}$, where $\lambda$ is a small
constant.
\item Repeat the previous step until all training examples are correctly
classified.
\end{enumerate}
If the data are linearly separable, the perceptron is guaranteed to find a
hyperplane that separates the two classes. However, it is not able to do this
when there is a non-linear decision boundary -- for that we will need the
multi-layer perceptron.

@@

\begin{figure}[h]
\label{fig:mlp}
\caption{}
\centering
%\includegraphics[scale=0.95]{images/a2-nn}
\end{figure}

@@

\section{Summary}
In this chapter we have outlined all of the learning algorithms that we will
be evaluating with respect to our classification problem, from the simple
0R and 1R to Na\"{i}ve Bayes, decision trees, logistic regression,
$k$-nearest neighbour (and extensions K* and Ranked Distance), to the more
sophisticated support vector machines and multi-layer perceptrons. The reason
for our inclusion of so many algorithms is two-fold. Firstly, many of
these algorithms have not been extensively applied in various medical settings
to the problem of LOS prediction. Using a wide range of algorithms from various
paradigms will enable us to see what algorithms perform better than others.
Secondly, evaluating several classifiers will allow us to make comparisons
between different feature selection methods that are independent of the
classifier and give us more insight into the nature of the data set and of the
problem.

>>>>

\chapter{Length of Stay Classification Using Supervised Learning}
  \label{chap:classification}

@@

\section{Zero-R and 1R}
\label{sec:zeror}
Short for ``0-rule'' and ``1-rule'' respectively, Zero-R and 1R are
very simple classifiers. Zero-R uses no features to predict the class, and
1R only uses one.

Given a data set, the Zero-R classification algorithm will find the class that
occurs the most often, and predict this class for all new examples. For LOS
classification, if one class in a data set was ``LOS $\leq$ 2 days'' and
the other was
``LOS $>$ 2 days'', and feature vectors (which represent patients in this
scenario) were more commonly labelled ``LOS $\leq$ 2 days'', then Zero-R will
predict all unseen patients with ``LOS $\leq$ 2 days''.
The prediction decision of Zero-R is very easy to understand,
and its results also provide an apt baseline for evaluating the more
sophisticated classifiers used in this work.

1R works by considering each feature individually.
For each of the values of the
feature, 1R finds the class with the highest frequency and creates a rule
from this observation. If the feature is numeric (that is, its values are
continuous rather than discrete), then 1R will partition the spectrum of
values into discrete buckets. This is known as \textit{discretisation}.
The error rate, expressed as the ratio of correct predictions to the total
number of examples in the data set,
is calculated for each rule that is generated
in this way. The rule with the lowest error rate is then chosen as the 1-rule
for this classifier,
and if there is a tie, it is broken randomly. As with Zero-R, the prediction
decision of 1R is transparent and it provides another baseline upon which
more sophisticated classifiers can be evaluated. Studies have also found that
1R performs quite well in terms of accuracy relative to other classifiers
\cite{Holte1993}.

Concretely, given a feature
vector of the form \textit{(age, gender, has\_cancer)},
1R will split the ages of
all patients into discrete categories
(such as \textit{0-20, 21-40, 41-65, 66+}). It
will then check the class labels for all training examples in the data set
that have \textit{age} in the 0-20 category
(as that is the first value for the discretised \textit{age} feature) and
note the most common class, proceeding similarly for the rest of the
categories of \textit{age}. Then it will consider \textit{gender}, which has
two possible values $\{$\textit{male, female}$\}$, and finally \textit{has\_cancer}
with two values $\{$\textit{true, false}$\}$. This results in 8 rules,
corresponding to three features.
An example of a rule is that \textit{age 0-20} most commonly
has ``LOS $\leq$ 2 days''. For each of these rules, 1R will attempt to
use only that rule to predict the LOS class for each example in the data set,
noting down the error rate. 1R then chooses the feature that produces rules
with the highest total accuracy.

\section{Na\"{i}ve Bayes}
Unlike 1R, which considers only one feature in the feature vector, Na\"{i}ve
Bayes (NB) classifiers allows all features to contribute equally to the
classification decision. Such classifiers apply Bayes' theorem with the
(unrealistic) assumption of independence between features, hence the
description \textit{na\"{i}ve}. This means that we
calculate the probability that a particular feature value belongs in a class
without considering the effect of other features on this probability. This
probability can be computed by
counting the frequency with which the value is associated with a class in the
training data. After these probabilities are computed, classifying a feature
vector simply involves multiplying the probabilities of each feature belonging
to each possible class. The predicted class for that feature vector is then the
class with the highest probability.

@@

Despite the na\"{i}ve independence assumption of the NB classifier, it often
works surprisingly well in practice and is relatively fast to train. We use it
in our work both as a simple classifier in its own right, and also as a more
sophisticated baseline to Zero-R and 1R.

\section{C4.5 Decision Tree}
\label{sec:c45}
Decision trees (DT) can be specified recursively and are an example of a
``divide-and-conquer'' approach to learning. First, we choose a feature to
place at the root node and then create one branch for each possible value.
This partitions the data set into subsets, one for every value that the
feature takes. We repeat this for each branch, choosing different features
each time and continuing to divide the data set into progressively smaller
subsets, growing the tree.
If all examples in a subset have the same class at any point during
this process, we stop splitting that subset.

@@

\section{Logistic Regression}
Logistic regression (LR) is a regression technique used for when the dependent
variable is dichotomous or two-valued, and thus it is a suitable candidate to
use for our LOS classification problem.
Previous work on trauma LOS prediction by
Dinh et. al. \cite{Dinh2013a}, as well as many attempts at modelling the LOS
in various medical domains, have all used logistic regression, often as a
starting point.

@@

Training the logistic regression classifier, then, is a matter of finding the
$\beta_i$ that will maximise the likelihood (conditional probability of the
data given the values for $\beta_i$) of the given data set. This is done using
numerical methods, where estimates of the $\beta_i$ are chosen, the likelihood
computed, and the $\beta_i$ are updated, and so on: this process is repeated
until parameter estimates converge or do not change more than a given threshold
value.

To classify a feature vector $\mathbf{x} = (x_1,x_2,\ldots,x_n)$, we substitute the
$x_i$ into the above equation in place of the $F_i$, which gives us a
probability. If this probability is above a particular threshold, then we
classify the example as belonging to the class ``LOS $\leq$ 2 days'', otherwise
it is classified as ``LOS $>$ 2 days''. Usually this threshold is 0.5.

\section{Support Vector Machine}
Support vector machines (SVMs) are an example of a supervised learning algorithm
capable of classifying training examples into two classes. They are one of the
best algorithms in terms of predictive accuracy \cite{Bellazzi2008}. Despite
this, they have not been widely applied to predicting the LOS, which gives us
an opportunity to evaluate their performance in our LOS classification problem.

@@

\begin{figure}[h]
\caption{An example of a 2-dimensional SVM.}
\label{fig:svm}
\centering
%\includegraphics[scale=0.95]{images/a2-nn}
\end{figure}

@@

\section{$k$-Nearest Neighbour}
In this section we will describe the $k$-nearest neighbour ($k$-NN) algorithm
and discuss some extensions which we use in our work. Additionally, we will
describe our contribution, which is an extension of the distance metric for
finding the nearest neighbours.

@@

\subsection{Existing and Proposed Extensions}
There have been many extensions proposed to the $k$-NN algorithm described
above. These usually seek to find a more suitable distance metric for the
problem domain at hand, or to reduce the storage requirements of the original
$k$-NN algorithm. We will focus on two extensions to the Euclidean distance:
K* and our proposed distance function, Ranked Distance.

@@

\subsubsection{Proposed Approach: Ranked Distance}
\label{sec:rankeddist}
\paragraph{Intuition}
As we mentioned earlier, all features are considered equally in a $k$-NN
classifier, which implies that all features are equally important in predicting
the class. In practice, however, this is not always the case: to predict the
LOS of a trauma patient, the severity of their injury should affect the final
LOS more than whether or not they can speak English (which could be just two
of many features that are recorded about a patient). Our contribution is thus
a modified distance function to take the relative importance of attributes into
account.

\paragraph{Mathematical formulation}
Given two instances $\mathbf{x} = (x_1,x_2,\ldots,x_n)$ and
$\mathbf{y} = (y_1,y_2,\ldots,y_n)$, the distance between them is given by:
\begin{equation*}
\mathrm{Distance} = \sum_{i=1}^n w_i |x_i-y_i|
\end{equation*}
where the $w_i$s weight the contribution of the $i$-th feature to the overall
distance that is computed by the $k$-NN algorithm. We assume that all feature
values have been normalised, as described above.

@@

\noindent We will use $f(\mathrm{Rank}) = \frac{1}{\mathrm{Rank}^c}$, where
$c$ varies from 0 to 1.
\end{enumerate}

\section{Multi-layer Perceptron}
Multi-layer perceptrons (MLPs) are a generalised form of the
\textit{perceptron}, first
studied extensively by Rosenblatt \cite{Rosenblatt1962}. The original
motivation of the perceptron 
was to develop a very simple artificial neural network to model the
behaviour of neurons in the brain. The perceptron model for a two-class
problem is a linear combination of the features: that is,
$h(\mathbf{x}) = \sum w_i x_i$ where the $w_i$ are the weights of the model.
There is an input node, called a \textit{neuron}, for each feature, and the
weights are applied to the connections between these input neurons to the
output.
Usually, a constant term of 1 called the \textit{bias} is also added to the
model for an additional trainable input. To classify an item, we substitute it
into $h(\mathbf{x})$; if the value is above a particular threshold, we label it
as class 1, else it belongs in class 2. For convenience, we can let this
threshold be 0. Training the perceptron involves the following steps:
\begin{enumerate}
\item Start with random numbers for the weights $w_i$.
\item For each training example $\mathbf{x_i}$: calculate $h(\mathbf{x_i})$.
If it is correctly classified, do not update the weights.
If it is incorrectly classified, add a multiple of the misclassified
instance to the current weights using
$\mathbf{w} = \mathbf{w} + \lambda \mathbf{x_i}$, where $\lambda$ is a small
constant.
\item Repeat the previous step until all training examples are correctly
classified.
\end{enumerate}
If the data are linearly separable, the perceptron is guaranteed to find a
hyperplane that separates the two classes. However, it is not able to do this
when there is a non-linear decision boundary -- for that we will need the
multi-layer perceptron.

@@

\begin{figure}[h]
\caption{An MLP with one hidden layer.}
\label{fig:mlp}
\centering
%\includegraphics[scale=0.95]{images/a2-nn}
\end{figure}

@@

\section{Summary}
In this chapter we have outlined all of the learning algorithms that we will
be evaluating with respect to our classification problem, from the simple
Zero-R and 1R to Na\"{i}ve Bayes, decision trees, logistic regression,
$k$-nearest neighbour (and extensions K* and Ranked Distance), to the more
sophisticated support vector machines and multi-layer perceptrons. The reason
for our inclusion of so many algorithms is two-fold. Firstly, many of
these algorithms have not been extensively applied in various medical settings
to the problem of LOS prediction. Using a wide range of algorithms from various
paradigms will enable us to see what algorithms perform better than others.
Secondly, evaluating several classifiers will allow us to make comparisons
between different feature selection methods that are independent of the
classifier and give us more insight into the nature of the data set and of the
problem.

### 2014-11-02 thesis/method/evaluation.tex
Fix broken table and figure labels in latex

\subsubsection{Repository}
All code relating to this work, including the \LaTeX code for this document,
as well as results from our evaluations are available publicly online at
\url{https://github.com/tianyupu/hons-thesis}. However, we were not able to
obtain permission from neither the Royal Prince Alfred Hospital nor the
Hospital das Foras Armadas to make the data publicly available due to the
privacy of patient information. Interested readers should discuss with the
author.

@@

\section{Performance Measures}
There are many ways to evaluate the performance of learning algorithms. The
two measures we will use are the \textit{accuracy} and the \textit{area under
the receiver operating characteristic (ROC) curve}, commonly abbreviated AUC
(for \textit{area under the curve}).

@@

The accuracy can also be expressed more simply and intuitively as
$\dfrac{\text{number of correct classifications}}{\text{total number of examples}}$
but the terms introduced above will be necessary to understand the
AUC measure below.

@@

\subsection{Area Under the Curve (AUC)}
The area under the receiver operating characteristic (ROC) curve, often
abbreviated AUC, is a measure of the ability of a binary classifier to
distinguish between members of the two classes. Let the \textit{true
positive rate} be $\dfrac{TP}{TP + FN}$ and the \textit{false positive rate}
be $\dfrac{FP}{FP + TN}$. The ROC curve is a plot of the true positive rate
on the vertical axis against the false positive rate on the horizontal axis.
The area under this curve will lie between 0 and 1, with 1 indicating a
100\% true positive rate and 0\% false positives (that is, everything was
classified correctly). The AUC of randomly picking between the two classes
with equal probability is 0.5, with a ROC curve that is a straight line
dividing the graph diagonally in half. Figure \ref{fig:sampleroc} shows
two ROC curves on the same graph:

\begin{figure}[h]
\label{fig:sampleroc}
\caption{}
\centering
%\includegraphics[scale=0.95]{images/a2-nn}
\end{figure}

@@

\subsection{Confidence Intervals}
In order to compare our results with that of Dinh et. al. \citep{Dinh2013a},
we are especially interested in the mean cross-validated AUC of our
classifiers as well as
the upper and lower bounds of the 95\% confidence interval associated with this
mean. The intervals are constructed using this argument:

For each classifier, let the population of all of its ten-fold cross-validated
AUC scores be some random variable $X$, which follows a distribution whose
parameters (such as mean and standard deviation) we do not know.
The Central Limit Theorem states that given a random
sample of $n$ items from $X$, namely $X_1,X_2,\dots,X_n$, the sample mean
$\bar{x}$
is approximately normally distributed when $n$ is at least roughly around 30
and the $X_i$'s are independent and identically distributed.
Consequently, we can construct a $(1-\alpha)$\% confidence interval for
$\bar{x}$ -- where $\alpha$ is commonly called the
\textit{significance level} -- as follows:
\begin{equation*}
  \left[\bar{x} - z_{1-\alpha/2}\dfrac{s}{\sqrt{n}},
    \bar{x} + z_{1-\alpha/2}\dfrac{s}{\sqrt{n}}\right]
\end{equation*}
where $\bar{x}$ is the sample mean, $\alpha$ is the significance level
(5\% or $0.05$ in our case), $s$ is the sample standard deviation, $n$
is the number of observations in our sample, and $z_{1-\alpha/2}$ is the value
of the standard normal variable $Z \sim \mathcal{N}(0,1)$ with cumulative
probability $1-\alpha/2$. Since we have taken 40 independent samples of the
ten-fold cross-validated AUC, the Central Limit Theorem holds and we can
compute the mean and standard deviation of our 40 samples in order to
construct a 95\% confidence interval for the mean AUC of each classifier.

>>>>

\subsubsection{Repository}
All code relating to this work, including the \LaTeX\ code for this document,
as well as results from our evaluations are available publicly online at
\url{https://github.com/tianyupu/hons-thesis}. However, we were not able to
obtain permission from neither the Royal Prince Alfred Hospital nor the
Hospital das Foras Armadas to make the data publicly available due to the
privacy of patient information. Interested readers should discuss with the
author.

@@

\section{Performance Measures}
There are many ways to evaluate the performance of learning algorithms. The
two measures we will use are the \textit{accuracy} and the \textit{area under
the receiver operating characteristic (ROC) curve}, commonly abbreviated AUC.

@@

The accuracy can also be expressed more simply and intuitively as
the number of correct classifications out of the total number of examples,
but the terms introduced above will be necessary to understand the
AUC measure below.

@@

\subsection{Area Under the Curve}
The area under the receiver operating characteristic (ROC) curve, often
abbreviated AUC, is a measure of the ability of a binary classifier to
distinguish between members of the two classes. Let the \textit{true
positive rate} be $\dfrac{TP}{TP + FN}$ and the \textit{false positive rate}
be $\dfrac{FP}{FP + TN}$. The ROC curve is a plot of the true positive rate
on the vertical axis against the false positive rate on the horizontal axis.
The area under this curve will lie between 0 and 1, with 1 indicating a
100\% true positive rate and 0\% false positives (that is, everything was
classified correctly). The AUC of randomly picking between the two classes
with equal probability is 0.5, with a ROC curve that is a straight line
dividing the graph diagonally in half. Figure \ref{fig:sampleroc} shows
two ROC curves on the same graph:

\begin{figure}[h]
\caption{Two ROC curves on the same graph. Note that the straight line from
the bottom left corner to the top right corner indicates an AUC of 0.5. Curves
which head towards the top left corner indicate increasing AUC.}
\label{fig:sampleroc}
\centering
%\includegraphics[scale=0.95]{images/a2-nn}
\end{figure}

@@

\subsection{Confidence Intervals}
In order to compare our results with our baseline, the work of Dinh et. al.
\citep{Dinh2013a}, as well as quantify the level of confidence we have in
our performance measures, we constructed 95\% confidence intervals for the
mean accuracy and AUC.

For each classifier, let the population of all of its accuracy (or AUC)
scores be some random variable $X$, which follows a distribution whose
parameters (such as mean and standard deviation) we do not know.
The Central Limit Theorem from statistics states that given a random
sample of $n$ items from $X$, namely $X_1,X_2,\dots,X_n$, the sample mean
$\bar{x}$
is approximately normally distributed when $n$ is at least roughly around 30
and the $X_i$'s are independent and identically distributed \cite{Phipps2001}.

Consequently, we can construct a $(1-\alpha)$\% normal confidence interval for
$\bar{x}$ -- where $\alpha$ is commonly called the
\textit{significance level} -- as follows \cite{Wasserman2003}:
\begin{equation*}
  \left[\bar{x} - z_{1-\alpha/2}\dfrac{s}{\sqrt{n}},
    \bar{x} + z_{1-\alpha/2}\dfrac{s}{\sqrt{n}}\right]
\end{equation*}
where $\bar{x}$ is the sample mean, $\alpha$ is the significance level
(5\% or $0.05$ in our case), $s$ is the sample standard deviation, $n$
is the number of observations in our sample, and $z_{1-\alpha/2}$ is the value
of the standard normal variable $Z \sim \mathcal{N}(0,1)$ with cumulative
probability $1-\alpha/2$.

These confidence intervals can be constructed

In our work, we repeated the ten-fold stratified cross-validation procedure
ten times, leading to $n=100$ samples with which to obtain our confidence
estimates.
